---
title: "Compound Interest Rates"
output:
  html_notebook: default
editor_options:
  chunk_output_type: console
---

The compound interest rate is the most powerful mathematical force in the financial world. It shows how money can grow inevitable and powerfully into the future...

First, lets create a function that describes compound interest:
```{r}
compound_int <- function(PV, Yeild, N) {
  PV * (1 + Yeild) ^ N  # compound interest formula
  }
```

### Examples to illustate compound interest
#### Feral Kittens
Lets say that we have 4 feral kittens terroizing your backyard. On an annual basis, the kitten population will grow by 50%. After 10 years, how many kittens will we have?
```{r}
compound_int(PV = 1, Yeild = 0.2, N = 35)
```

230 is a lot a kittens...

#### Snowball example
We have a quarter pound snowball that starts rolling down Mt. Everest. This snowball gains weight at rate of 10% for every 10 yards. How heavy will the snowball be after rolling 1 mile down and falling off a cliff?

```{r}
# There are 1,760 yars per mile, growth occurs every 10 yards, 
# so it rolls 176 yards. 

compound_int(PV = 0.25, Yeild = 0.10, N = 176)
```

Phew, 4.8 million pounds after 1 mile, starting from 1/4 pound. Point being you can start small and end up very, very big. 

#### Real money
So, if we started with $1, continually invested for 500 years with a 5% growth rate would yeild:
```{r}
compound_int(PV = 1, Yeild = 0.05, N = 500)
```
Whoa, 39 BILLION, with a B. 

These extreme examples and assumptions illustrate the extremes that compound growth can get to. 


### Real examples

#### Growth on my IRA
Lets say im 30, and have 5,000 I can contribute to my IRA account. What will that 5,000 look like when I retire at 70?
```{r}
PV = 5000  # initial contribution
Yeild = 0.10  # interest rate or return
N = 40  # number of compounding periods
compound_int(PV = PV, Yeild = Yeild, N = N)
```

The future value of my 5,000 would be 108,622. Not bad. And if you also contributed and additional 5,000 every year for 10-15 years, I could be a millionaire in old age. 

### Takeaways
The compound interest formula can explain the growth of inflation as in rental costs, to government debt, to credit card payments over time. Or, how modest contribitions to a retirement account at an early age can guarantee we become millionaires. 

## More-often-than-annual-compounding
Many things compound more than anually:
+ Bonds - 2/yr
+ Dividend stocks - 4/yr
+ Ferral kittens - 6/yr 
+ credit card debt - 12/yr

So, we must make our formula a little more complicated:
PV = principal amount
Yeild = interest rate/yeild/return
Time = Time period in years
N = number of compounding periods per year

```{r}
compound_int_2 <- function(PV, Yeild, N, Time) {
  PV * (1 + (Yeild / N)) ^ (N * Time)  # compound interest formula
  }
```

So lets compare simple annual compounding:
```{r}
compound_int(PV = 50, Yeild = 0.04, N = 40)
```

With more-than-annual compounding:
```{r}
compound_int_2(PV = 50, Yeild = 0.04, N = 4, Time = 40)
```

Its ~2.3% larger than with simple annual compounding. Point being, this is why bonds and divident accounts are great, and how credit card companies screw us...